---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Downloading the data.

```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl, destfile = "./repdata%2Fdata%2Factivity.zip", method = "curl")
```

Unzipping the data.

```{r}
unzip("repdata%2Fdata%2Factivity.zip")
```

Loading the data set.

```{r}
activity <- read.csv("activity.csv")
```

Saving the dataset as data frame.

```{r}
activity <- data.frame(activity)
str(activity)
```

## What is mean total number of steps taken per day?

Calculating the number of steps taken per day.

```{r}
activity_date <-aggregate(steps ~ date, activity, sum)
str(activity_date)
```

Histogram of the total number of steps taken each day.

```{r histogram, fig.height=5.5}
hist(activity_date$steps, xlab = "Steps per Day",
     main = "Total Number of Steps", col = "blue")
```

Calculating the mean and median of the total number of steps taken per day.

```{r}
mean(activity_date$steps, na.rm = TRUE)
median(activity_date$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
