---
title: "Reproducible Research Project 1"
author: "_Teo Moreno_"
date: "_16 Oct 2015_"
output: html_document
---


### Loading and preprocessing the data

Loading the dataset.

```{r}
activity <- read.csv("activity.csv")
```

Saving the dataset as data frame.

```{r}
activity <- data.frame(activity)
str(activity)
```

### What is the total number of steps taken per day?

Calculating the number of steps taken per day.

```{r}
activity_date <- aggregate(steps ~ date, activity, sum)
str(activity_date)
```

Histogram of the total number of steps taken each day.

```{r histogram, fig.height = 5.5}
hist(activity_date$steps, xlab = "Steps per Day",
     main = "Total Number of Steps", col = "blue")
```

Calculating the mean and median of the total number of steps taken per day.

```{r}
mean(activity_date$steps, na.rm = TRUE)
median(activity_date$steps, na.rm = TRUE)
```

### What is the average daily activity pattern?

Calculating the average daily activity by 5-minute interval.

```{r}
activity_interval <- aggregate(steps ~ interval, activity, mean)
str(activity_interval)
```

Plotting steps versus the 5-minute interval.

```{r plot, fig.height=5.5}
plot(activity_interval$interval, activity_interval$steps,
     xlab= 'Interval', ylab= 'Steps', type = "l",
     main = "Activity Pattern", col = "blue")
```

Calculating the 5-minute interval that contains the maximum number of steps.

```{r}
activity_interval$interval[which.max(activity_interval$steps)]
```

### Imputing missing values

Calculating and reporting the total number of missing values in the dataset.

```{r}
sum(!complete.cases(activity))
```

Filling all the missing values in the dataset with the mean across intervals.

```{r}
activity_na <- activity[which(!complete.cases(activity)), ]
activity_na$steps <- mean(activity_interval$steps, na.rm = TRUE)
str(activity_na)
```

Creating a new dataset combining complete data an filled missing data.

```{r}
activity_comp <- activity[which(complete.cases(activity)), ]
str(activity_comp)
activity_fill <- rbind(activity_comp, activity_na)
index <- order(activity_fill$date)
activity_fill <- activity_fill[index, ]
str(activity_fill)
```

New calculatiion of the number of steps taken per day.

```{r}
activity_fill_date <-aggregate(steps ~ date, activity_fill, sum)
str(activity_fill_date)
```

New Histogram of the total number of steps taken each day.

```{r histogram2, fig.height=5.5}
hist(activity_fill_date$steps, xlab = "Steps per Day",
     main = "New Total Number of Steps", col = "blue")
```

New mean and median of the total number of steps taken per day.

```{r}
mean(activity_fill_date$steps)
median(activity_fill_date$steps)
```

### Are there differences in activity patterns between weekdays and weekends?

Adding a weekday column to the activity set.

```{r}
activity_fill$weekday <- weekdays(as.Date(activity_fill$date))
```

Adding a daytype column to the activity set.

```{r}
activity_fill$daytype <- ifelse(activity_fill$weekday == "Saturday" | 
                                       activity_fill$weekday == "Sunday", "Weekend", "Weekday") 
```

Calculating the average activity by day type (weekday or weekend).

```{r}
activity_fill_daytype <- aggregate(steps~interval+daytype, activity_fill, mean)
str(activity_fill_daytype)
```

Making a time series plot.

```{r plot2, fig.height=7.5}
library(lattice)
xyplot(steps~interval|daytype, activity_fill_daytype, layout = c(1,2),
       type = "l", main = "Activity Pattern")
```